---
title: "The Sharpe Ratio"
description: |
  In this post we present a classic finance use case using the PerformanceAnalytics, quantmod, and dygraphs packages. We'll demonstrate importing stock data, building a portfolio, and then calculating the Sharpe Ratio. 
preview: preview.png
author: 
  - name: "Jonathan Regenstein"
    url: https://www.linkedin.com/in/jkregenstein/
    affiliation: RStudio
    affiliation_url: https://www.rstudio.com
date: 11-09-2016
output: radix::radix_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# packages required for this post
for (pkg in c('PerformanceAnalytics', 'quantmod', 'dygraphs', 'RColorBrewer')) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)
}
```

## Overview

In this post we will demonstrate the calculation of a Sharpe Ratio for a stock portfolio. We'll start with a function that grabs monthly stock returns and saves those monthly returns as an xts object in the global environment. With that function, we will create three xts objects of monthly returns, and merge those three xts objects into one object, before passing that merged object to dygraphs to peek at the individual stocks. 

Then, we'll move on to build a portfolio by selecting asset weights and calculating the portfolio monthly returns. Next, we will calculate the growth of a dollar invested in that portfolio (which is what matters to us) over time, and save the results to an xts object. Dygraphs will come in handy again for the portfolio visualizations. Finally, we will calculate the Sharpe Ratio.

## Setting up

We start by loading three packages: quantmod to download the data, PerformanceAnalytics to run portfolio calculations, and dygraphs to graph time series objects. We will also create a function to import stock data.

```{r message=FALSE}
library(PerformanceAnalytics)
library(quantmod)
library(dygraphs)

# Function to calculate monthly returns on a stock
monthly_stock_returns <- function(ticker, start_year) {
  
  # Download the data from Yahoo finance
  symbol <- getSymbols(ticker, src = 'yahoo', 
                       auto.assign = FALSE, warnings = FALSE)
  
  # Tranform it to monthly returns using quantmode::periodReturn
  data <- periodReturn(symbol, period = 'monthly', 
                       subset=paste(start_year, "::", sep = ""),
                       type = 'log')

  # Let's rename the column of returns to something intuitive because
  # the column name is what will eventually be displayed
  colnames(data) <- as.character(ticker)

  # We want to be able to work with the xts objects so let's explicitly
  # assign them into the global environment using ticker name 
  assign(ticker, data, .GlobalEnv)
}
```

The `monthly_stock_returns` function above takes 2 parameters, a stock symbol and a year. Note that we could have included a third parameter called something like ‘period' if we wanted the ability to grab periods other than monthly returns. For example, we can envision a desire to look at annual, weekly or daily returns. Here, I force monthly returns because I don't want to allow different period options. It's a choice driven by the purpose of this Notebook - which here is focused on monthly returns.

## Preparing the data

In the next chunk, we choose three stock tickers and a starting year argument for the `monthly_stock_returns` function. Then, we merge them into one xts object.

```{r, cache=TRUE}
# Choose the starting year and assign it to the 'year' variable
year <- 2010

# Use the function the monthly returns on 3 stocks, and pass in the 'year'
# value. Let's choose Google, JP Morgan and Amazon
monthly_stock_returns('GOOG', year)
monthly_stock_returns('JPM', year)
monthly_stock_returns('AMZN', year)

# Merge the 3 monthly return xts objects into 1 xts object.
merged_returns <- merge.xts(GOOG, JPM, AMZN)
```

Let's graph the individual performances of each stock over time.

```{r, layout="l-page"}
# Before we combine these into a portfolio, graph the individual returns
# and see if anything jumps out as unusual. It looks like something 
# affected Google in March of 2014, but didn't affect JP Morgan or Amazon.
dygraph(merged_returns, main = "Google v JP Morgan v Amazon") %>%
  dyAxis("y", label = "%") %>%
  dyOptions(colors = RColorBrewer::brewer.pal(3, "Set2"))
```

Nothing earth-shattering thus far: we have an xts object of three time series and have seen that one of them had weird behavior in April of 2014 (a Google stock split). We'll ignore that behavior for this example and go on to constructing a portfolio.

## Calculating portfolio returns

Here we'll find the monthly returns of a weighted combination of assets. Unsurprisingly, we start out by choosing those weights.

```{r, layout="l-page"}
# We have the 3 monthly returns saved in 1 object.
# Now, let's choose the respective weights of those 3.
# Here we'll allocate 25% to Google, 25% to JP Morgan and 50% to Amazon.
w <- c(.25, .25, .50)

# Now use the built in PerformanceAnalytics function Return.portfolio
# to calculate the monthly returns on the portfolio,
portfolio_monthly_returns <- Return.portfolio(merged_returns, weights = w)

# Use dygraphs to chart the portfolio monthly returns.
dygraph(portfolio_monthly_returns, main = "Portfolio Monthly Return") %>%
  dyAxis("y", label = "%")
```

Now, instead of looking at monthly returns, let's look at how $1 would have grown in this portfolio.

```{r, layout="l-page"}
# Add the wealth.index = TRUE argument and, instead of monthly returns,
# the function will return the growth of $1 invested in the portfolio.
dollar_growth <- Return.portfolio(merged_returns, weights = w, 
                                  wealth.index = TRUE)

# Use dygraphs to chart the growth of $1 in the portfolio.
dygraph(dollar_growth, main = "Growth of $1 Invested in Portfolio") %>%
  dyAxis("y", label = "$")
```

A dollar would have grown quite nicely in this portfolio - about 2.5x - fantastic. 

## Calculating the Sharpe Ratio

Now let's look at the risk/reward of this portfolio by calculating the Sharpe Ratio. Briefly, the Sharpe Ratio is the mean of the excess monthly returns above the risk-free rate, divided by the standard deviation of the excess monthly returns above the risk-free rate. This is the formulation of the Sharpe Ratio as of 1994; if we wished to use the original formulation from 1966 the denominator would be the standard deviation of portfolio monthly returns. Learn more [here](http://web.stanford.edu/~wfsharpe/art/sr/sr.htm).

In other words, the Sharpe Ratio measures excess returns per unit of volatility, where we take the standard deviation to represent portfolio volatility. The Sharpe Ratio was brought to us by Bill Sharpe - arguably the most important economist for modern investment management as the creator of the Sharpe Ratio, CAPM and Financial Engines, a forerunner of today's robo-advisor movement.

In the code chunk below, we'll calculate the Sharpe Ratio in two ways.

First, we'll use the Return.excess function from PerformanceAnalytics to calculate a time series of monthly excess returns. Two arguments need to be supplied: the time series of returns and the risk-free rate. The function will return a time series of excess returns, and we'll take the mean of that time series to get the numerator of the Sharpe Ratio. Then we'll divide by the standard deviation of the that time series to get the Sharpe Ratio.

Our second method is a bit easier. We'll use the SharpeRatio function in PerformanceAnalytics, for which we'll supply two arguments: a time series of monthly returns and risk-free rate.

For both methods I use a risk-free rate of .03% as the approximate mean of the 1-month Treasury bill rate since 2010. I'll cover a quick way to grab this and other data via Quandl in a future post.

```{r}
# Method 1: use the Return.excess function from PerformanceAnalytics,
# then calculate the Sharpe Ratio manually.
portfolio_excess_returns <- Return.excess(portfolio_monthly_returns, 
                                          Rf = .0003)
sharpe_ratio_manual <- round(
  mean(portfolio_excess_returns) / StdDev(portfolio_excess_returns), 4
)

# If we wanted to use the original, 1966 formulation of the Sharpe Ratio,
# there is one small change to the code in Method 1
sharpe_ratio <- round(
  SharpeRatio(portfolio_monthly_returns, Rf = .0003), 4
)
```

Using the Return.excess function and then dividing by the standard deviation of excess returns, the Sharpe Ratio is `sharpe_ratio_manual[,1]` = `r sharpe_ratio_manual[,1]`.

Using the built in SharpeRatio function, the Sharpe Ratio is `sharpe_ratio[1,]` = `r sharpe_ratio[1,]`.

Alright, we have built a portfolio and calculated the Sharpe Ratio - and also set up some nice reusable chunks for data import, portfolio construction and visualization. We haven't done anything terribly complex but this can serve as a useful paradigm to any collaborators, including our future selves, who want to reproduce this work, learn from this work, or expand upon this work.

