---
title: "Radix for R Markdown"
description: Scientific and technical writing, native to the web
author:
  - name: "JJ Allaire"
    url: https://github.com/jjallaire
    affiliation: RStudio
    affiliation_url: https://www.rstudio.com
  - name: "Rich Iannone"
    url: https://github.com/rich-iannone
    affiliation: RStudio
    affiliation_url: https://www.rstudio.com
  - name: "Yihui Xie"
    url: https://github.com/yihui
    affiliation: RStudio
    affiliation_url: https://www.rstudio.com
date: September 10, 2018
citation_url: https://rstudio.github.io/radix
bibliography: biblio.bib
output:
  radix::radix_article:
    toc: false
site: radix::radix_website
---


```{r setup, include=FALSE}
library(ggplot2)
library(knitr)
```

Radix is an online publication format designed for scientific and technical communication. Radix articles include:

- Reader-friendly typography that adapts well to mobile devices.
- Features essential to technical writing like [LaTeX math](#equations), [citations], and [footnotes](#footnotes-and-asides).
- Flexible [figure layout](figures.html) options (e.g. displaying figures at a larger width than the article text).
- Attractively rendered [tables](tables.html) with optional support for pagination.
- Support for a wide variety of [diagramming tools](diagrams.html) for illustrating concepts.
- The ability to incorporate JavaScript and D3-based [interactive visualizations](interactivity.html).
- A variety of ways to [publish articles](publish_article.html), including support for publishing sets of articles as a [Radix website](website.html).
- The ability to [create a blog](blog.html) composed of a collection of Radix articles.


Radix is based on the [Distill web framework](https://github.com/distillpub/template), which was originally created for use in the Distill Machine Learning Journal [@distill]. Radix combines the technical authoring features of Distill with [R Markdown](https://rmarkdown.rstudio.com/), enabling a fully reproducible workflow based on literate programming [@knuth1984].

## Installation

To create an [R Markdown](https://rmarkdown.rstudio.com) document that uses the Radix format, first install the **radix** R package from GitHub:

```r
devtools::install_github("rstudio/radix")
```

Using Radix requires Pandoc v2.0 or higher. If you are using RStudio then you should use RStudio v1.2.718 or higher (which comes bundled with Pandoc v2.0). 

You can download the preview release of RStudio v1.2 at <https://www.rstudio.com/products/rstudio/download/preview/>.

```{r, child="_common/radix_basics.Rmd"}
```




