---
title: "Distill for R Markdown"
description: Scientific and technical writing, native to the web
author:
  - name: "JJ Allaire"
    url: https://github.com/jjallaire
    affiliation: Posit Software, PBC
    affiliation_url: https://www.posit.co
    orcid_id: 0000-0003-0174-9868
  - name: "Rich Iannone"
    url: https://github.com/rich-iannone
    affiliation: Posit Software, PBC
    affiliation_url: https://www.posit.co
  - first_name: "Alison"
    last_name: "Presmanes Hill"
    affiliation: Posit Software, PBC
    affiliation_url: https://www.posit.co
    orcid_id: 0000-0002-8082-1890
  - name: "Yihui Xie"
    url: https://github.com/yihui
    affiliation: Posit Software, PBC
    affiliation_url: https://www.posit.co
    orcid_id: 0000-0003-0645-5666
  - name: "Christophe Dervieux"
    url: https://github.com/cderv
    affiliation: Posit Software, PBC
    affiliation_url: https://www.posit.co
    orcid_id: 0000-0003-0645-5666

date: September 10, 2018
citation_url: https://rstudio.github.io/distill
bibliography: biblio.bib
link-citations: true
output:
  distill::distill_article:
    toc: true
site: distill::distill_website
---

```{r setup, include=FALSE}
library(ggplot2)
library(knitr)
```

Distill for R Markdown is a web publishing format optimized for scientific and technical communication. Distill articles include:

-   Reader-friendly typography that adapts well to mobile devices.
-   Features essential to technical writing like [LaTeX math](#equations), [citations], and [footnotes](#footnotes-and-asides).
-   Flexible [figure layout](figures.html) options (e.g. displaying figures at a larger width than the article text).
-   Attractively rendered [tables](tables.html) with optional support for pagination.
-   Support for a wide variety of [diagramming tools](diagrams.html) for illustrating concepts.
-   The ability to incorporate JavaScript and D3-based [interactive visualizations](interactivity.html).
-   A variety of ways to [publish articles](publish_article.html), including support for publishing sets of articles as a [Distill website](website.html) or as a [Distill blog](blog.html).

Distill for R Markdown is based on the [Distill web framework](https://github.com/distillpub/template), which was originally created for use in the Distill Machine Learning Journal [@distill]. Distill for R Markdown combines the technical authoring features of Distill with [R Markdown](https://rmarkdown.rstudio.com/), enabling a fully reproducible workflow based on literate programming [@knuth1984].

```{r, child="_common/distill_basics.Rmd"}
```
