---
title: "JavaScript & D3"
description: Enhance communication with interactive visualizations 
preview: images/javascript-d3-preview.png
---

There are a variety of ways to add interactive JavaScript and D3 visualizations to your Distill articles, including:

-   Using pre-built [htmlwidgets](http://www.htmlwidgets.org) that wrap JavaScript visualization libraries like [Leaflet](http://rstudio.github.io/leaflet/), [Plotly](https://plot.ly/r/), [dygraphs](http://rstudio.github.io/dygraphs/), and [threejs](https://bwlewis.github.io/rthreejs/).

-   Creating custom [D3.js](https://d3js.org/) interactive visualizations using the [r2d3](https://rstudio.github.io/r2d3/) package.

## htmlwidgets

Including [htmlwidgets](http://www.htmlwidgets.org) within a Distill article is as easy as including an R plot. For example, the following code embeds a Leaflet map:

```` clike
```{r}`r ''`
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
```
````

```{r}
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
```

You can also use [figure layout](figures.html) options with htmlwidgets. For example, here we set `layout="l-page"` to specify that we want the widget to span the width of the page:

```` clike
```{r, layout="l-page", fig.height=3}`r ''`
library(dygraphs)
dygraph(ldeaths)
```
````

```{r, layout="l-page", fig.height=3}
library(dygraphs)
dygraph(ldeaths)
```

Note that we specify `fig.height=3` to provide a more natural height for a widget that spans the entire width of the page.

To learn about available htmlwidgets see the [showcase page](http://www.htmlwidgets.org/showcase_leaflet.html) and the [htmlwidget gallery](http://gallery.htmlwidgets.org/).

## D3 visualizations

The [r2d3](https://rstudio.github.io/r2d3/) package enables you to easily incorporate interactive graphics created using [D3.js](https://d3js.org/) in your Distill articles.

For example, the following renders an interactive [Voronoi diagram](https://en.wikipedia.org/wiki/Voronoi_diagram) created using D3:

```` clike
```{r, layout="l-page", fig.height=6}`r ''`
library(r2d3)
r2d3(script = "voronoi.js", d3_version = 4)
```
````

```{r, layout="l-page", fig.height=6}
library(r2d3)
r2d3(script = "d3/voronoi/voronoi.js", d3_version = 4)
```

Note that we specify `layout="l-page"` to indicate that we want the visualization to occupy the full width of the page.
