---
title: "Interactivity"
description: Enhance communication with interactive visualizations 
---

There are a variety of ways to add interactivity to your Distill articles, including:

- Using pre-built [htmlwidgets](http://www.htmlwidgets.org) that wrap JavaScript visualization libraries like [Leaflet](http://rstudio.github.io/leaflet/), [Plotly](https://plot.ly/r/), [dygraphs](http://rstudio.github.io/dygraphs/), and [threejs](https://bwlewis.github.io/rthreejs/).

- Creating custom [D3.js](https://d3js.org/) interactive visualizations using the [r2d3](https://rstudio.github.io/r2d3/) package.

- Using [Shiny](https://shiny.rstudio.com) to create an [interactive document](https://rmarkdown.rstudio.com/authoring_shiny.html#embedded_applications) that embeds interactive components within your article.

## htmlwidgets

Including [htmlwidgets](http://www.htmlwidgets.org) within a Distill article is as easy as including an R plot. For example, the following code embeds a Leaflet map:

````clike
```{r}`r ''`
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
```
````

```{r}
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
```

You can also use [custom layout] with htmlwidgets. For example, here we set `distill.layout="l-page"` to specify that we want the widget to span the width of the page:

````clike
```{r, distill.layout="l-page", fig.height=2}`r ''`
library(dygraphs)
dygraph(ldeaths) %>% 
  dyRangeSelector()
```
````

```{r, distill.layout="l-page", fig.height=2}
library(dygraphs)
dygraph(ldeaths) %>% 
  dyRangeSelector()
```

Note that we specify `fig.height=2` to provide a more natural height for a widget that spans the entire width of the page.

To learn about available htmlwidgets see the [showcase page](http://www.htmlwidgets.org/showcase_leaflet.html) and the [htmlwidget gallery](http://gallery.htmlwidgets.org/).

