---
title: "Tables"
description: Display data in attractive, compact formats
---

There are a number of options available for displaying tables within Distill articles, including:

-   Using the `knitr::kable()` function to render a data frame as HTML.

-   Using the `rmarkdown::paged_table()` function to create a page-able version of a data frame.

-   Use the [gtsummary](http://www.danieldsjoberg.com/gtsummary/) and [gt](https://gt.rstudio.com/) packages to create to create publication-ready analytical and summary tables.

-   Using standard [markdown table syntax](https://pandoc.org/MANUAL.html#tables) to create a custom table.

## Knitr kable

The `knitr::kable()` function will render an R data frame as an HTML table. For example:

```` {.clike}
```{r, layout="l-body-outset"}`r ''`
library(knitr)
kable(head(mtcars))
```
````

```{r, layout="l-body-outset"}
library(knitr)
kable(head(mtcars))
```

Often times tables will require more width for their display than the standard Distill article text width. Here we use `layout="l-body-outset"` to cause the table to outset slightly from the main text column. All of the available [figure layouts](figures.html) can also be used with tables.

## Paged tables

If a data frame has a large number of rows it might not be practical to display it fully inline. In these situations you can use the `rmarkdown::paged_table()` function to create a table that allows pagination of rows and columns. For example:

```` {.clike}
```{r, layout="l-body-outset"}`r ''`
library(rmarkdown)
paged_table(mtcars)
```
````

```{r, layout="l-body-outset"}
library(rmarkdown)
paged_table(mtcars)
```

There are a number of options available that control the display of paged tables. By default, 10 rows at a time are displayed but you can use the `rows.print` option to increase this. For example:

```` {.clike}
```{r, layout="l-body-outset"}`r ''`
library(rmarkdown)
paged_table(mtcars, options = list(rows.print = 15))
```
````

```{r, layout="l-body-outset"}
library(rmarkdown)
paged_table(mtcars, options = list(rows.print = 15, cols.print = 5))
```

Here is a summary of the available paged table options:

| Option           | Description                                    |
|------------------|------------------------------------------------|
| `rows.print`     | Maximum rows to print per page.                |
| `max.print`      | Maximum rows in the table (defaults to 1000).  |
| `cols.print`     | Maximum columns in the table (defaults to 10). |
| `rownames.print` | Print row names as part of the table.          |

## gtsummary tables

The [gtsummary](http://www.danieldsjoberg.com/gtsummary/) package provides an elegant and flexible way to create publication-ready analytical and summary tables. The package summarizes data sets, regression models, and more, using sensible defaults with highly customizable capabilities. For example:

```` {.clike}
```{r}`r ''`
library(gtsummary)
library(survival)

# build glm table
mod1 <- glm(response ~ trt + age + grade, trial, family = binomial)
t1 <- tbl_regression(mod1, exponentiate = TRUE)

# build survival model table
t2 <-
  coxph(Surv(ttdeath, death) ~ trt + grade + age, trial) %>%
  tbl_regression(exponentiate = TRUE)

# merge tables 
tbl_merge(
  tbls = list(t1, t2),
  tab_spanner = c("**Tumor Response**", "**Time to Death**")
)
```
````

```{r}
library(gtsummary)
library(survival)

# build glm table
mod1 <- glm(response ~ trt + age + grade, trial, family = binomial)
t1 <- tbl_regression(mod1, exponentiate = TRUE)

# build survival model table
t2 <-
  coxph(Surv(ttdeath, death) ~ trt + grade + age, trial) %>%
  tbl_regression(exponentiate = TRUE)

# merge tables 
tbl_merge(
  tbls = list(t1, t2),
  tab_spanner = c("**Tumor Response**", "**Time to Death**")
)
```

## gt tables

The [gtsummary](http://www.danieldsjoberg.com/gtsummary/) package is built in top of the [gt](https://gt.rstudio.com/) package, which provides a flexible foundation for creating just about any sort of table you can imagine. Here is a brief example of how to use **gt** to create a table from the included `sp500` dataset:

```` {.clike}
```{r}`r ''`
library(gt)
library(tidyverse)
library(glue)

# Define the start and end dates for the data range
start_date <- "2010-06-07"
end_date <- "2010-06-14"

# Create a gt table based on preprocessed
# `sp500` table data
sp500 %>%
  dplyr::filter(date >= start_date & date <= end_date) %>%
  dplyr::select(-adj_close) %>%
  gt() %>%
  tab_header(
    title = "S&P 500",
    subtitle = glue::glue("{start_date} to {end_date}")
  ) %>%
  fmt_date(
    columns = vars(date),
    date_style = 3
  ) %>%
  fmt_currency(
    columns = vars(open, high, low, close),
    currency = "USD"
  ) %>%
  fmt_number(
    columns = vars(volume),
    suffixing = TRUE
  )
```
````

```{r}
library(gt)
library(tidyverse)
library(glue)

# Define the start and end dates for the data range
start_date <- "2010-06-07"
end_date <- "2010-06-14"

# Create a gt table based on preprocessed
# `sp500` table data
sp500 %>%
  dplyr::filter(date >= start_date & date <= end_date) %>%
  dplyr::select(-adj_close) %>%
  gt() %>%
  tab_header(
    title = "S&P 500",
    subtitle = glue::glue("{start_date} to {end_date}")
  ) %>%
  fmt_date(
    columns = vars(date),
    date_style = 3
  ) %>%
  fmt_currency(
    columns = vars(open, high, low, close),
    currency = "USD"
  ) %>%
  fmt_number(
    columns = vars(volume),
    suffixing = TRUE
  )
```

## Markdown tables

You can also define [tables using markdown](https://pandoc.org/MANUAL.html#tables). For example, the table from the previous section that describes paged table options was defined with the following markdown:

``` {.markdown}
| Option           | Description                                    |
|------------------|------------------------------------------------|
| `rows.print`     | Maximum rows to print per page.                |
| `max.print`      | Maximum rows in the table (defaults to 1000).  |
| `cols.print`     | Maximum columns in the table (defaults to 10). |
| `rownames.print` | Print row names as part of the table.          |
```

You can use Distill [figure layout](figures.html) classes with markdown tables by enclosing them in markdown div (`:::`) named with the appropriate class, for example:

``` {.markdown}
::: l-body-outset
| Option           | Description                                    |
|------------------|------------------------------------------------|
| `rows.print`     | Maximum rows to print per page.                |
| `max.print`      | Maximum rows in the table (defaults to 1000).  |
| `cols.print`     | Maximum columns in the table (defaults to 10). |
| `rownames.print` | Print row names as part of the table.          |
:::
```
