---
title: "Tables"
description: Display data in attractive, compact formats
---

There are a number of options available for displaying tables within Distill articles, including:

- Using the `knitr::kable()` function to render a data frame as HTML.

- Using the `rmarkdown::paged_table()` function to create a page-able version of a data frame.

- Using standard [markdown table syntax](https://pandoc.org/MANUAL.html#tables) to create a custom table.

## Knitr kable

The `knitr::kable()` function will render an R data frame as an HTML table. For example:

````clike
```{r, layout="l-body-outset"}`r ''`
library(knitr)
kable(head(mtcars))
```
````


```{r, layout="l-body-outset"}
library(knitr)
kable(head(mtcars))
```

Often times tables will require more width for their display than the standard Distill article text width. Here we use `layout="l-body-outset"` to cause the table to outset slightly from the main text column. All of the available [figure layouts](figures.html) can also be used with tables.

## Paged tables

If a data frame has a large number of rows it might not be practical to display it fully inline. In these situations you can use the `rmarkdown::paged_table()` function to create a table that allows pagination of rows and columns. For example:

````clike
```{r, layout="l-body-outset"}`r ''`
library(rmarkdown)
paged_table(mtcars)
```
````

```{r, layout="l-body-outset"}
library(rmarkdown)
paged_table(mtcars)
```

There are a number of options available that control the display of paged tables. By default, 10 rows at a time are displayed but you can use the `rows.print` option to increase this. For example:

````clike
```{r, layout="l-body-outset"}`r ''`
library(rmarkdown)
paged_table(mtcars, options = list(rows.print = 15))
```
````

```{r, layout="l-body-outset"}
library(rmarkdown)
paged_table(mtcars, options = list(rows.print = 15, cols.print = 5))
```

Here is a summary of the available paged table options:

| Option           | Description                                    |
|------------------|------------------------------------------------|
| `rows.print`     | Maximum rows to print per page.                |
| `max.print`      | Maximum rows in the table (defaults to 1000).  |
| `cols.print`     | Maximum columns in the table (defaults to 10). |
| `rownames.print` | Print row names as part of the table.          |

## Markdown tables

You can also define [tables using markdown](https://pandoc.org/MANUAL.html#tables). For example, the table from the previous section that describes paged table options was defined with the following markdown:

```markdown
| Option           | Description                                    |
|------------------|------------------------------------------------|
| `rows.print`     | Maximum rows to print per page.                |
| `max.print`      | Maximum rows in the table (defaults to 1000).  |
| `cols.print`     | Maximum columns in the table (defaults to 10). |
| `rownames.print` | Print row names as part of the table.          |
```

You can use Distill [figure layout](figures.html) classes with markdown tables by enclosing them in markdown div (`:::`) named with the appropriate class, for example:

```markdown
::: l-body-outset
| Option           | Description                                    |
|------------------|------------------------------------------------|
| `rows.print`     | Maximum rows to print per page.                |
| `max.print`      | Maximum rows in the table (defaults to 1000).  |
| `cols.print`     | Maximum columns in the table (defaults to 10). |
| `rownames.print` | Print row names as part of the table.          |
:::
```






